---
title: "AET/PET ratios for BGC units"
author: "Hardy"
date: "11/20/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message=FALSE)

# Load libraries
library(tidyverse)
library(forestDroughtTool)
library(lubridate)
library(magrittr)

# Load some functions
source("C:/Users/hgriesba/Documents/git/forestDroughtTool/R/asmrCalc.R")
source("C:/Users/hgriesba/Documents/git/forestDroughtTool/R/yearSelect.R")


```

```{r define cleanData function}
# imputeTS has now changed so have to recode function

cleanData<-function(climData) {
  
  # Omit years with 10 or more consecutive missing values in any climate variable

  climData %>%
  # filter(stn==stnID) %>% 
  group_by(year) %>%
  summarise(pptNA=imputeTS::statsNA(ppt,print_only=FALSE)$longest_na_gap,
            tmxNA=imputeTS::statsNA(tmx,print_only=FALSE)$longest_na_gap,
            tmnNA=imputeTS::statsNA(tmn,print_only=FALSE)$longest_na_gap)%>%
  mutate(pptNA, pptNA = ifelse(is.na(pptNA), 0, pptNA)) %>%
  mutate(tmxNA, tmxNA = ifelse(is.na(tmxNA), 0, tmxNA)) %>%
  mutate(tmnNA, tmnNA = ifelse(is.na(tmnNA), 0, tmnNA)) %>%
  filter(pptNA<10&tmxNA<10&tmnNA<10) %>%
  select(year) %>%
  
  # Impute NA values using adjacent values
  left_join(climData,by="year") %>%
  mutate(ppt2=imputeTS::na_locf(ppt),ppt3=imputeTS::na_locf(ppt,option="nocb")) %>%
  mutate(tmx2=imputeTS::na_locf(tmx),tmx3=imputeTS::na_locf(tmx,option="nocb")) %>%
  mutate(tmn2=imputeTS::na_locf(tmn),tmn3=imputeTS::na_locf(tmn,option="nocb")) %>%
  rowwise() %>%
  mutate(ppt_filled=mean(ppt2,ppt3)) %>%
  mutate(tmx_filled=mean(tmx2,tmx3)) %>%
  mutate(tmn_filled=mean(tmn2,tmn3)) %>%
  dplyr::select(year,month,day,ppt_filled,tmx_filled,tmn_filled) %>% 
 
     return()
}



```

```{r import stn data,message=FALSE,warning=FALSE,eval=FALSE}

# This won't run, takes a long time

# Compile climate data into a list
X<-list(

    apex<-
      readxl::read_excel("ApexRoadside_daily.xlsx") %>% 
      mutate(year=year(Day),
             month=month(Day),
             day=day(Day)) %>% 
      rename(tmn="Tmin",
             tmx="Tmax",
             ppt="Precip") %>% 
     mutate(stn="apex") %>% 
      dplyr::select(stn,year,month,day,tmn,tmx,ppt) %>% 
      mutate_at(vars(year:ppt),as.numeric), 
      
    
    beaverdell<-
      readxl::read_excel("Beaverdell_daily.xlsx") %>% 
      mutate(year=year(Day),
             month=month(Day),
             day=day(Day)) %>% 
      rename(tmn="Tmin",
             tmx="Tmax",
             ppt="Precip") %>% 
    mutate(stn="beaverdell") %>% 
      dplyr::select(stn,year,month,day,tmn,tmx,ppt)%>% 
      mutate_at(vars(year:ppt),as.numeric),
    
    bigCreek<-
      readxl::read_excel("BigCreek_1893-1998.xlsx",range="B2:J38686",skip=1,col_types = "numeric") %>% 
      transmute(stn="bigCreek",
                year=Year,
             month=Month,
             day=Day,
             tmn=Tmin,
             tmx=Tmax,
             ppt=Total)%>% 
      mutate_at(vars(year:ppt),as.numeric),
    
    bmn<-
      readxl::read_excel("BMN_1993-2020.xlsx",skip=1) %>% 
      setNames(c("time","ppt","tmx","tmn")) %>% 
      mutate(stn="bmn",
             year=year(time),
             month=month(time),
             day=day(time)) %>% 
      dplyr::select(stn,year,month,day,tmn,tmx,ppt)%>% 
      mutate_at(vars(year:ppt),as.numeric),
    
    
    fording<-
      readxl::read_excel("FordingRiverCominco_1970-2017.xlsx",skip=0) %>% 
        transmute(stn="fording",
                year=Year,
             month=Month,
             day=Day,
             tmn=Tmin,
             tmx=Tmax,
             ppt=Total)%>% 
      mutate_at(vars(year:ppt),as.numeric),
    
    greenstone<-
      readxl::read_excel("GreenstoneHub_daily.xlsx") %>%   
      mutate(year=year(Day),
             month=month(Day),
             day=day(Day)) %>% 
      rename(tmn="Tmin",
             tmx="Tmax",
             ppt="Precip") %>% 
    mutate(stn="greenstone") %>% 
      dplyr::select(stn,year,month,day,tmn,tmx,ppt)%>% 
      mutate_at(vars(year:ppt),as.numeric),
    
    idaBell<-
      readxl::read_excel("IdaBell3_daily.xlsx") %>%   
      mutate(year=year(Day),
             month=month(Day),
             day=day(Day)) %>% 
      rename(tmn="Tmin",
             tmx="Tmax",
             ppt="Precip") %>% 
    mutate(stn="idaBell") %>% 
      dplyr::select(stn,year,month,day,tmn,tmx,ppt)%>% 
      mutate_at(vars(year:ppt),as.numeric),
    
    jellicoe<-
      readxl::read_excel("JellicoeDaily_1995-2020.xlsx") %>% 
        transmute(stn="jellicoe",
                year=Year,
             month=Month,
             day=Day,
             tmn=Tmin,
             tmx=Tmax,
             ppt=Total)%>% 
      mutate_at(vars(year:ppt),as.numeric),
    
    kimberley<-
      readxl::read_excel("KimberleyPCC_1975-2020.xlsx") %>% 
        transmute(stn="kimberley",
                year=Year,
             month=Month,
             day=Day,
             tmn=Tmin,
             tmx=Tmax,
             ppt=Total)%>% 
      mutate_at(vars(year:ppt),as.numeric),
    
    pennask <- 
      readxl::read_excel("PennaskSummit_daily.xlsx") %>%   
      mutate(year=year(Day),
             month=month(Day),
             day=day(Day)) %>% 
      rename(tmn="Tmin",
             tmx="Tmax",
             ppt="Precip") %>% 
    mutate(stn="pennask") %>% 
      dplyr::select(stn,year,month,day,tmn,tmx,ppt)%>% 
      mutate_at(vars(year:ppt),as.numeric),
    
    redLake<-
        readxl::read_excel("RedLakeDaily_1974-2020.xlsx") %>% 
        transmute(stn="redLake",
                year=Year,
             month=Month,
             day=Day,
             tmn=Tmin,
             tmx=Tmax,
             ppt=Total)%>% 
      mutate_at(vars(year:ppt),as.numeric),
    
    sparwood<-
        readxl::read_excel("Sparwood_1980-2020.xlsx") %>% 
        transmute(stn="sparwood",
                year=Year,
             month=Month,
             day=Day,
             tmn=Tmin,
             tmx=Tmax,
             ppt=Total)%>% 
      mutate_at(vars(year:ppt),as.numeric),
    
    wasa<-
       readxl::read_excel("Wasa_1924-2017.xlsx",skip=1,col_types = "numeric") %>% 
      transmute(stn="wasa",
                year=Year,
             month=Month,
             day=Day,
             tmn=Tmin...6,
             tmx=Tmax...5,
             ppt=Total...10)%>% 
      mutate_at(vars(year:ppt),as.numeric)
  
) # close list
  
# now rbind the data frames
rawData<-
  bind_rows(X) %>% 
  mutate(stn=factor(stn)) %>% 
  ungroup()

# apply cleanData function
cleanedData<-  
  rawData %>% 
  group_by(stn) %>% 
  group_modify(~ cleanData(.x)) %>% 
  ungroup()

climData<-
  list(rawData=rawData,cleanedData=cleanedData)

save(climData,file="climData.RData")

```

Here is table summarizing the number of years for each station that have sufficient climate data for AET/PET calculation, as per DeLong et al. (2019):

```{r summarize data}

# Load climate data (compiled from previous step)
load("climData.RData")

# Summarize data
climData %>% 
  pluck("cleanedData") %>% 
  rename(Station="stn") %>% 
  mutate(Station=str_to_title(Station)) %>% 
  group_by(Station) %>% 
  summarise(NumYrs=length(unique(year)),
            From=min(year),
            To=max(year)) %>% 
  knitr::kable()



```

## BGC unit ASMR   

```{r function to summarise ASMR, echo=FALSE}

asmrRun1<-function(stnID,bgc) {

  # set seed to control random year 
  set.seed(5)  
  
    # Generate random years
  asmrData<-
    climData %>% 
    pluck("cleanedData") %>%
    filter(stn==stnID) %>% 
    rename(ppt="ppt_filled",
           tmx="tmx_filled",
           tmn="tmn_filled") %>% 
    mutate(date=paste(year,month,day,sep="-")) %>% 
    asmrCalc()
  
  yearData<-
    asmrData %>% 
    yearSelect(excl=c(1970,2020),win=c(1980,2005),yrs=10)
  

  asmrSummary<- 
   asmrData %>% 
    filter(year %in% yearData) %>% # filter for years
    group_by(month) %>% 
    summarise_at(vars(ends_with(".ASMR")),mean) %>%
        ungroup() %>%
    
        rbind(c(13,colMeans(.)[-1])) %>%
        mutate(month=cut(month,breaks=13,labels=c(month.abb,"Annual"))) %>% 
    mutate(Station=str_to_title(stnID),BGC=bgc) %>% 
    mutate_at(vars(ends_with(".ASMR")),round,3) %>% 
    dplyr::select(Station,BGC,everything()) 
  
  return(list(years=yearData,asmr=asmrSummary))
 

  
  
}
  


```

For each BGC unit, I chose the station with the best climate data coverage.  we can run this again if there's a better station to use.

---------------

## IDFdm1

```{r calculate ASMR_}
stnID="kimberley"
bgc="IDFdm1"

x<-asmrRun1(stnID,bgc)

```

For this BGC unit, I used `r stnID` station, and randomly selected the following years: `r x$years`.  Here is the monthly and annual AET/PET ratio for those 10 years:

```{r show output_}

x$asmr %>% 
  knitr::kable()

```


---------------------

## MSdm1 and MSdm2

```{r calculate ASMR1}
stnID="sparwood"
bgc="MSdm1"

x<-asmrRun1(stnID,bgc)

```

For this BGC unit, I used `r stnID %>% str_to_title()` station, and randomly selected the following years: `r x$years`.  Here is the monthly and annual AET/PET ratio for those 10 years:

```{r show output1}

x$asmr %>% 
  knitr::kable()
```

---------------------

## MSxk1

```{r calculate ASMR2}
stnID="apex"
bgc="MSxk1"

x<-asmrRun1(stnID,bgc)

```

For this BGC unit, I used `r stnID %>% str_to_title()` station, and randomly selected the following years: `r x$years`.  Here is the monthly and annual AET/PET ratio for those 10 years:

```{r show output2}

x$asmr %>% 
  knitr::kable()
```

-------------------

## IDFdk1

```{r calculate ASMR3}
stnID="redLake"
bgc="IDFdk1"

x<-asmrRun1(stnID,bgc)

```

For this BGC unit, I used `r stnID %>% str_to_title()` station, and randomly selected the following years: `r x$years`.  Here is the monthly and annual AET/PET ratio for those 10 years:

```{r show output3}

x$asmr %>% 
  knitr::kable()
```

-------------------

## IDFdk2

```{r calculate ASMR4}
stnID="jellicoe"
bgc="IDFdk2"

x<-asmrRun1(stnID,bgc)

```

For this BGC unit, I used `r stnID %>% str_to_title()` station, and randomly selected the following years: `r x$years`.  Here is the monthly and annual AET/PET ratio for those 10 years:

```{r show output4}

x$asmr %>% 
  knitr::kable()
```
-------------------

## ESSFdc2

```{r calculate ASMR5}
stnID="pennask"
bgc="ESSFdc2"

x<-asmrRun1(stnID,bgc)

```

For this BGC unit, I used `r stnID %>% str_to_title()` station, and randomly selected the following years: `r x$years`.  Here is the monthly and annual AET/PET ratio for those 10 years:

```{r show output5}

x$asmr %>% 
  knitr::kable()
```

-------------------

## ESSFxc2
Greenstone Hub is one of the stations, but only has 4 years of cleaned data.  Is there another station to use?  The spreadsheet says "Mission Creek" but I didn't see the data there.
