---
title: "AET/PET ratios for BGC units"
author: "Hardy"
date: "11/20/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message=FALSE)

# Load libraries
library(tidyverse)
library(forestDroughtTool)
library(lubridate)
library(magrittr)
library(here)

# Load some functions
# source("C:/Users/hgriesba/Documents/git/forestDroughtTool/R/asmrCalc.R")
# source("C:/Users/hgriesba/Documents/git/forestDroughtTool/R/yearSelect.R")


# Load some functions
source("C:/Users/hgriesba/Documents/git/bcgov/forestDroughtTool/R/asmrCalc.R")
source("C:/Users/hgriesba/Documents/git/bcgov/forestDroughtTool/R/yearSelect.R")

```

```{r define cleanData function}
# imputeTS has now changed so have to recode function

cleanData<-function(climData) {
  
  # Omit years with 10 or more consecutive missing values in any climate variable

  climData %>%
  # filter(stn==stnID) %>% 
  group_by(year) %>%
  summarise(pptNA=imputeTS::statsNA(ppt,print_only=FALSE)$longest_na_gap,
            tmxNA=imputeTS::statsNA(tmx,print_only=FALSE)$longest_na_gap,
            tmnNA=imputeTS::statsNA(tmn,print_only=FALSE)$longest_na_gap)%>%
  mutate(pptNA, pptNA = ifelse(is.na(pptNA), 0, pptNA)) %>%
  mutate(tmxNA, tmxNA = ifelse(is.na(tmxNA), 0, tmxNA)) %>%
  mutate(tmnNA, tmnNA = ifelse(is.na(tmnNA), 0, tmnNA)) %>%
  filter(pptNA<10&tmxNA<10&tmnNA<10) %>%
  select(year) %>%
  
  # Impute NA values using adjacent values
  left_join(climData,by="year") %>%
  mutate(ppt2=imputeTS::na_locf(ppt),ppt3=imputeTS::na_locf(ppt,option="nocb")) %>%
  mutate(tmx2=imputeTS::na_locf(tmx),tmx3=imputeTS::na_locf(tmx,option="nocb")) %>%
  mutate(tmn2=imputeTS::na_locf(tmn),tmn3=imputeTS::na_locf(tmn,option="nocb")) %>%
  rowwise() %>%
  mutate(ppt_filled=mean(ppt2,ppt3)) %>%
  mutate(tmx_filled=mean(tmx2,tmx3)) %>%
  mutate(tmn_filled=mean(tmn2,tmn3)) %>%
  dplyr::select(year,month,day,ppt_filled,tmx_filled,tmn_filled) %>% 
 
     return()
}



```

## Clean IDFxx data

```{r clean IDFxx data}

x<-readRDS(here::here("data-raw","WxData_newBECvariants","IDFxx","IDFxx_data.rds"))

# grab Cranbrook data
cran.Data<-
  x %>% 
  filter(station_name=="CRANBROOK A") %>% 
  mutate(year=as.numeric(year)) %>% 
  filter(year>1968) %>% 
  dplyr::select(stn=station_name,
                year,month,day,
                tmn=min_temp,
                tmx=max_temp,
                ppt=total_precip) %>% 
  cleanData() %>% 
  mutate(Station="Cranbrook") %>% 
  dplyr::select(Station,everything()) %>% 
  
  # break into two datasets
  mutate(Station=replace(Station,year>=1991,"Cranbrook2"))


# grab Wardner data
ward.Data<-
  x %>% 
  filter(station_name=="WARDNER KTNY HATCHERY") %>% 
    mutate(year=as.numeric(year)) %>% 
  filter(year>1971) %>% 
  dplyr::select(stn=station_name,
                year,month,day,
                tmn=min_temp,
                tmx=max_temp,
                ppt=total_precip) %>% 
  cleanData %>% 
  mutate(Station="Wardner") %>% 
  dplyr::select(Station,everything())%>% 
  
  # break into two datasets
  mutate(Station=replace(Station,year>=1991,"Wardner2"))



cleanedData<-
 
  rbind(ward.Data,cran.Data) %>% 
   as.data.frame() %>% 
  mutate(month=as.numeric(month),
         day=as.numeric(day))
  


```





Here is table summarizing the number of years for each station that have sufficient climate data for AET/PET calculation, as per DeLong et al. (2019):

```{r summarize data}


# Summarize data
cleanedData %>% 
   
group_by(Station) %>% 
  summarise(NumYrs=length(unique(year)),
            From=min(year),
            To=max(year)) %>% 
  knitr::kable()




```

```{r create future dataset,eval=FALSE}

load(here("data","bgcClim.rda"))

clim.X<-
  bgcClim %>% 
  pivot_longer(cols=Tmax01:last_col()) %>% 
  mutate(Period=paste(period,Scenario,sep="_")) %>% 
  dplyr::select(-period,-Scenario) %>% 
  pivot_wider(names_from="Period",values_from="value")



future.ppt<-
  clim.X %>% 
  dplyr::filter(str_detect(name,"PPT")) %>% 
  mutate_at(vars(matches("_rcp")), list(dif = (~ . / `1961 - 1990_NA`)))

future.tmx<-
  clim.X %>% 
  dplyr::filter(str_detect(name,"Tmax")) %>% 
  mutate_at(vars(matches("_rcp")), list(dif = (~ . - `1961 - 1990_NA`)))


future.tmn<-
  clim.X %>% 
  dplyr::filter(str_detect(name,"Tmin")) %>% 
  mutate_at(vars(matches("_rcp")), list(dif = (~ . - `1961 - 1990_NA`)))


# Future climate data
  future.Data<-
    rbind(future.ppt,future.tmx,future.tmn) %>% 
    mutate(month=str_sub(name,start=-2,end=-1)) %>% 
    mutate(climVar=str_sub(name,start=1,end=-3)) %>% 
    dplyr::select(BGC,climVar,month,contains("dif")) %>% 
    pivot_longer(`2025_rcp45_dif`:last_col(),names_to="scenario",values_to="values") %>% 
    mutate(climVar=paste(climVar,"future",sep="_")) %>% 
    pivot_wider(names_from="climVar",values_from="values") %>% 
    mutate(month=as.integer(month))

save(future.Data,file=here::here("data-raw","WxData_newBECvariants","futureData.RData"))

```

```{r load future data}

load(here::here("data-raw","WxData_newBECvariants","futureData.RData"))

```


## BGC unit ASMR   

```{r function to summarise ASMR, echo=FALSE}

asmrRun1<-function(stnID,bgc) {

  # set seed to control random year 
  set.seed(5)  
  
    # Generate random years
  asmrData<-
    cleanedData %>% 
    as.data.frame() %>% 
    mutate(Station=factor(Station)) %>% 
     # .[.$Station%in%stnID,] %>% # for some reason dplyr::filter isn't working
    filter(Station %in% stnID) %>% 
    rename(ppt="ppt_filled",
           tmx="tmx_filled",
           tmn="tmn_filled") %>% 
    mutate(date=paste(year,month,day,sep="-")) %>% 
    asmrCalc()
  
  
  yearData<-
    asmrData %>% 
    yearSelect(excl=c(1960,2020),win=c(1970,2020),yrs=10)
  

  
  asmrSummary<- 
   asmrData %>% 
    filter(year %in% yearData) %>% # filter for years
    group_by(month) %>% 
    summarise_at(vars(ends_with(".ASMR")),mean) %>%
        ungroup() %>%
    
        rbind(c(13,colMeans(.)[-1])) %>%
        mutate(month=cut(month,breaks=13,labels=c(month.abb,"Annual"))) %>% 
    mutate(Station=str_to_title(stnID),BGC=bgc) %>% 
    mutate_at(vars(ends_with(".ASMR")),round,3) %>% 
    dplyr::select(Station,BGC,everything()) 
  
  
  
    # generate future AET/PET ratios
 future.X<-
   cleanedData %>% 
   as.data.frame() %>% 
    filter(Station %in% stnID) %>% 
    rename(ppt="ppt_filled",
           tmx="tmx_filled",
           tmn="tmn_filled") %>% 
   mutate_at(vars("year","month","day"),as.numeric) %>% 
    right_join(filter(future.Data,BGC==bgc),by="month") %>% 
    mutate(ppt=PPT_future*ppt,
           tmx=Tmax_future+tmx,
           tmn=Tmin_future+tmn) %>% 
   mutate(date=paste(year,month,day,sep="-")) %>% 
   dplyr::select(Station,BGC,year,month,day,date,ppt,tmx,tmn,scenario)
    
   futureASMR<-rbind(data.frame(Scenario="rcp45_2025",asmrCalc(filter(future.X,scenario=="2025_rcp45_dif"))),
    data.frame(Scenario="rcp45_2055",asmrCalc(filter(future.X,scenario=="2055_rcp45_dif"))),
    data.frame(Scenario="rcp85_2025",asmrCalc(filter(future.X,scenario=="2025_rcp85_dif"))),
    data.frame(Scenario="rcp85_2055",asmrCalc(filter(future.X,scenario=="2055_rcp85_dif")))) %>% 
   
   # create summaries
       filter(year %in% yearData) %>% # filter for years
    group_by(Scenario,month) %>% 
    summarise_at(vars(ends_with(".ASMR")),mean) %>%
        ungroup() %>%
        mutate(Station=str_to_title(stnID),BGC=bgc) %>% 
    mutate_at(vars(ends_with(".ASMR")),round,3) %>% 
     mutate(month=as.integer(month)) %>% 
    dplyr::select(Station,BGC,everything()) 
     
   
  futureSummary<- 
     futureASMR %>% 
       mutate(month=as.integer(month)) %>% 
      ungroup() %>% 
     group_by(Scenario) %>% 
     summarise_at(vars(ends_with(".ASMR")),mean) %>% 
        mutate(Station=str_to_title(stnID),BGC=bgc,month=13) %>%
           dplyr::select(Station,BGC,month,everything()) %>% 
       rbind(futureASMR) %>% 
          arrange(Scenario,month) %>% 
         mutate(month=cut(month,breaks=13,labels=c(month.abb,"Annual"))) %>% 
    mutate_at(vars(ends_with(".ASMR")),round,3) %>% 
   mutate(Period=str_split_fixed(Scenario,patter="_",n=2)[,2]) %>% 
     mutate(Scenario=str_split_fixed(Scenario,patter="_",n=2)[,1]) %>% 
    dplyr::select(-Station) %>% 
    dplyr::select(BGC,Period,Scenario,Month=month,everything())
 
  
  
  return(list(years=yearData,asmr=asmrSummary,future=futureSummary))
 

  
  
}
  


```



-----------------

## IDFxx2 using Cranbrook data

### 1961-1990 period

```{r calculate ASMR_marysville}
stnID="Cranbrook"
bgc="IDFxx2"

x<-asmrRun1(stnID,bgc)

```

The following years were randomly selected: `r x$years`.  Here is the monthly and annual AET/PET ratio for those 10 years:

```{r show output_marysville}

x$asmr %>% 
  knitr::kable()

```


Future data are summarized in the table below:

```{r}

x$future %>% 
  knitr::kable()

```

---------------

### 1991-2020 period

```{r calculate ASMR_}
stnID="Cranbrook2"
bgc="IDFxx2"

x<-asmrRun1(stnID,bgc)

```

The following years were randomly selected: `r x$years`.  Here is the monthly and annual AET/PET ratio for those 10 years:

```{r show output_}

x$asmr %>% 
  knitr::kable()

```


Future data are summarized in the table below:

```{r}

x$future %>% 
  knitr::kable()

```

---------------------

## IDFxx2 using Wardner data

### 1961-1990 period

```{r}
stnID="Wardner"
bgc="IDFxx2"

x<-asmrRun1(stnID,bgc)

```

The following years were randomly selected: `r x$years`.  Here is the monthly and annual AET/PET ratio for those 10 years:

```{r show output_wardner}

x$asmr %>% 
  knitr::kable()

```


Future data are summarized in the table below:

```{r}

x$future %>% 
  knitr::kable()

```

---------------

### 1991-2020 period

```{r calculate ASMR_wardner2}
stnID="Wardner2"
bgc="IDFxx2"

x<-asmrRun1(stnID,bgc)

```

The following years were randomly selected: `r x$years`.  Here is the monthly and annual AET/PET ratio for those 10 years:

```{r show output_wardner2}

x$asmr %>% 
  knitr::kable()

```


Future data are summarized in the table below:

```{r}

x$future %>% 
  knitr::kable()

```

---------------------
